---
title: "State Map Development"
author: "Evan Canfield"
date: "9/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library Calls
The following libraries were used to develop the map.
```{r library calls, message=FALSE, warning=FALSE}
if (!require(pacman)) {install.packages('pacman')} 
p_load(
    broom
  , mapproj
  , maptools
  , noncensus
  , rgeos
  , rgdal
  , RColorBrewer
  , socviz
  , scales
  , stringr
  , tidyverse
  , viridis
)
```

# Read in the GeoJSON Files
The U.S. state-level maps in the Socviz library were prepared using shapefiles from the U.S. Census Bureau that were converted to GeoJSON format by [Eric Celeste](https://eric.clst.org/tech/usgeojson/). I used the 5m file. If this source is no longer available raw shapefiles can be obtained from the [US Census Bureau](https://www.census.gov/programs-surveys/geography.html). Shape files can then be transformed to GeoJSON via free online converters, such as [MyGeodata Converter](https://mygeodata.cloud/converter).
```{r }
us_states <- readOGR(
    dsn = "./data/gz_2010_us_040_00_5m.json"
    )

us_counties <- readOGR(
    dsn = "./data/gz_2010_us_050_00_5m.json"
    )
```
# State Level Map
## Transform Coordinate System
```{r}
us_states_aea <- spTransform(us_states,
                    CRS("+proj=laea +lat_0=45 +lon_0=-100 \
                         +x_0=0 +y_0=0 +a=6370997 +b=6370997 \
                         +units=m +no_defs"))

# Convert Row Names to Column
us_states_aea@data$id <- rownames(us_states_aea@data)
```

## Re-size and Shift States
### Alaska
```{r}
# Isolate Alaska Portion of the Shapefile
alaska_s <- us_states_aea[us_states_aea$STATE == "02",]

# Rotate Alaska 50 degrees
alaska_s <- elide(alaska_s, rotate=-50)

# Rescale Aaska
alaska_s <- elide(alaska_s, scale=max(apply(bbox(alaska_s), 1, diff)) / 2.3)

# Shift Alaska Coordiantes to Below California / Arizona
alaska_s <- elide(alaska_s, shift=c(-2.3e6, -2.475e6))

# Retrieve Projection Information
proj4string(alaska_s) <- proj4string(us_states_aea)
```

### Hawaii
```{r}
# Isolate Hawaii Portion of the Shapefile
hawaii_s <- us_states_aea[us_states_aea$STATE=="15",]

# Rotate Hawaii 35 degrees
hawaii_s <- elide(hawaii_s, rotate=-35)

# Shift Hawaii Coordiantes to Below Arizona / New Mexico / Texas
hawaii_s <- elide(hawaii_s, shift=c(5.4e6, -1.4e6))

# Retrieve Projection Information
proj4string(hawaii_s) <- proj4string(us_states_aea)
```

## Recombine Shifted Files
Note that Puerto Rico will be dropped in its entirety. If you have Puerto Rico data Healy recommends to shift it to between Texas and Florida
```{r}
# Drop Original Alaska, Hawaii, and Puerto Rico
us_states_aea <- us_states_aea[!us_states_aea$STATE %in% c("02", "15", "72"),]

# Bind Alaska and Hawaii
us_states_aea <- rbind(us_states_aea, alaska_s, hawaii_s)
```

## Tidy Shapefile
The united states shape-file needs to be converted into a data frame for easy use with ggplot. The State FIPS code is also created here. Once created the ID code is no longer necessary and is dropped.
```{r warning=FALSE}
us_state_map  <- tidy(us_states_aea, region = "GEO_ID") %>% 
  mutate(state_fips = str_sub(id,-2,-1)) %>% 
  select(-id)
```

## Export State-Level Data Frame

The final state-level data frame is saved as file *us_map.RDS*.
```{r}
saveRDS(
    object = us_state_map
  , file = "./us_state_map.RDS"
  )
```

# County Level Map
## Transform Coordinate System
```{r}
us_county_aea <- spTransform(us_counties,
                    CRS("+proj=laea +lat_0=45 +lon_0=-100 \
                         +x_0=0 +y_0=0 +a=6370997 +b=6370997 \
                         +units=m +no_defs"))

# Convert Row Names to Column
us_county_aea@data$id <- rownames(us_county_aea@data)
```

## Re-size and Shift States
### Alaska
```{r}
# Isolate Alaska Portion of the Shapefile
alaska_c <- us_county_aea[us_county_aea$STATE == "02",]

# Rotate Alaska 50 degrees
alaska_c <- elide(alaska_c, rotate=-50)

# Rescale Aaska
alaska_c <- elide(alaska_c, scale=max(apply(bbox(alaska_c), 1, diff)) / 2.3)

# Shift Alaska Coordiantes to Below California / Arizona
alaska_c <- elide(alaska_c, shift=c(-2.3e6, -2.475e6))

# Retrieve Projection Information
proj4string(alaska_c) <- proj4string(us_county_aea)
```

### Hawaii
```{r}
# Isolate Hawaii Portion of the Shapefile
hawaii_c <- us_county_aea[us_county_aea$STATE=="15",]

# Rotate Hawaii 35 degrees
hawaii_c <- elide(hawaii_c, rotate=-35)

# Shift Hawaii Coordiantes to Below Arizona / New Mexico / Texas
hawaii_c <- elide(hawaii_c, shift=c(5.4e6, -1.4e6))

# Retrieve Projection Information
proj4string(hawaii_c) <- proj4string(us_county_aea)
```

## Recombine Shifted Files
Note that Puerto Rico will be dropped in its entirety. If you have Puerto Rico data Healy recommends to shift it to between Texas and Florida
```{r}
# Drop Original Alaska, Hawaii, and Puerto Rico
us_county_aea <- us_county_aea[!us_county_aea$STATE %in% c("02", "15", "72"),]

# Bind Alaska and Hawaii
us_county_aea <- rbind(us_county_aea, alaska_c, hawaii_c)
```

## Tidy Shapefile
The united states shape-file needs to be converted into a data frame for easy use with ggplot. The State FIPS code is also created here. Once created the ID code is no longer necessary and is dropped.
```{r warning=FALSE}
us_county_map  <- tidy(us_county_aea, region = "GEO_ID") %>% 
  mutate(
    state_fips = str_sub(id,-5,-4),
    county_fips = str_sub(id,-3,-1),
    fips = str_sub(id,-5,-1)
    ) %>% 
  select(-id)
```

## Export County-Level Data Frame
The final state-level data frame is saved as file *us_map.RDS*.
```{r}
saveRDS(
    object = us_county_map
  , file = "./us_county_map.RDS"
  )
```

# Testing
In order to test that the state and county maps display correctly and can connect to data based on the proper FIPS code, data from the socviz package will be used. 

## Creating Data Frames
```{r}

# Create List of Travel Time By Complete FIPS code
counties_travel_time <- socviz::county_data %>% 
  select(id, travel_time)

# Merge with Spatial Data
us_county_map_travel <- us_county_map %>% 
  left_join(counties_travel_time, by = c("fips" = "id"))
```



### Inset Lines on GGPlot
I also developed lines to be places on a ggplot map to create boxes around Alaska and Hawaii. I followed the same aesthetic format as Figure 7.1 in Healy. 
## County Map - No Inset Boxes
```{r}
c <- ggplot()

c1 <- c + geom_polygon(
    data = us_county_map_travel 
  , mapping = aes(x = long, y = lat, group = group, fill = travel_time)
  , color = "gray90"
  , alpha = 0.75
  , size = 0.005) + 
  coord_equal() +
  scale_fill_viridis_c(direction = -1) +
  guides(fill=guide_legend(title="Travel Time"))+
  labs(title = "Average Daily Travel Time By County") +
  theme_void()

c1
```

## County Map - With Inset Boxes
```{r}
# Line Properties
line_color = "black"
line_size = 0.5

c2 <- c1 + 
  geom_segment(aes(x = -2.10e6, y = -1.50e6, xend = -1.15e6, yend = -1.50e6)
               , color =  line_color, size = line_size
               ) + 
  geom_segment(aes(x = -1.15e6, y = -1.50e6, xend = -7.75e5, yend = -1.85e6)
               , color =  line_color, size = line_size
               ) +
  geom_segment(aes(x = -7.75e5, y = -1.85e6, xend = -7.75e5, yend = -2.20e6)
               , color =  line_color, size = line_size
               ) +
  geom_segment(aes(x = -7.75e5, y = -1.85e6, xend = -2.90e5, yend = -1.85e6)
               , color =  line_color, size = line_size
               ) +
  geom_segment(aes(x = -2.90e5, y = -1.85e6, xend =  1.00e5, yend = -2.20e6)
               , color =  line_color, size = line_size
               ) +
  geom_segment(aes(x =  1.00e5, y = -2.20e6, xend =  1.00e5, yend = -2.40e6)
               , color =  line_color, size = line_size
               ) +
  geom_segment(aes(x = -7.75e5, y = -2.20e6, xend = -5.60e5, yend = -2.40e6)
               , color =  line_color, size = line_size
               )

c2
```

## County Map With State Borders - No Inset Boxes
```{r}
s_c <-  ggplot()

s_c1 <- s_c + 
  geom_polygon(
    data = us_county_map_travel 
  , mapping = aes(x = long, y = lat, group = group, fill = travel_time)
  , color = "gray90"
  , alpha = 0.75
  , size = 0.25) +
    geom_polygon(
    data = us_state_map
  , mapping = aes(x = long, y = lat, group = group)
  , color = "#1A5276"
  , alpha = 0
  , size = 1)
  
s_c2 <- s_c1 + 
  coord_equal() +
  scale_fill_viridis_c(direction = -1) +
  guides(fill=guide_legend(title="Travel Time"))+
  labs(title = "Average Daily Travel Time By County") +
  theme_void()

s_c2
```

## County Map With State Borders - With Inset Boxes
```{r}
# Line Properties
line_color = "black"
line_size = 0.5

s_c3 <- s_c2 + 
  geom_segment(aes(x = -2.10e6, y = -1.50e6, xend = -1.15e6, yend = -1.50e6)
               , color =  line_color, size = line_size
               ) + 
  geom_segment(aes(x = -1.15e6, y = -1.50e6, xend = -7.75e5, yend = -1.85e6)
               , color =  line_color, size = line_size
               ) +
  geom_segment(aes(x = -7.75e5, y = -1.85e6, xend = -7.75e5, yend = -2.20e6)
               , color =  line_color, size = line_size
               ) +
  geom_segment(aes(x = -7.75e5, y = -1.85e6, xend = -2.90e5, yend = -1.85e6)
               , color =  line_color, size = line_size
               ) +
  geom_segment(aes(x = -2.90e5, y = -1.85e6, xend =  1.00e5, yend = -2.20e6)
               , color =  line_color, size = line_size
               ) +
  geom_segment(aes(x =  1.00e5, y = -2.20e6, xend =  1.00e5, yend = -2.40e6)
               , color =  line_color, size = line_size
               ) +
  geom_segment(aes(x = -7.75e5, y = -2.20e6, xend = -5.60e5, yend = -2.40e6)
               , color =  line_color, size = line_size
               )

s_c3
```