---
title: "State Map Development"
author: "Evan Canfield"
date: "9/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library Calls
The following libraries were used to develop the map.
```{r library calls, message=FALSE, warning=FALSE}
if (!require(pacman)) {install.packages('pacman')} 
p_load(
    broom
  , mapproj
  , maptools
  , rgeos
  , rgdal
  , socviz
  , scales
  , stringr
  , tidyverse
  , viridis
)
```

# Read in the GeoJSON File
The U.S. state-level maps in the Socviz library were prepared using shapefiles from the U.S. Census Bureau that were converted to GeoJSON format by [Eric Celeste](https://eric.clst.org/tech/usgeojson/). I used the 5m file. If this source is no longer available raw shapefiles can be obtained from the [US Census Bureau](https://www.census.gov/programs-surveys/geography.html). Shape files can then be transformed to GeoJSON via free online converters, such as [MyGeodata Converter](https://mygeodata.cloud/converter).
```{r }
us_states <- readOGR(
    dsn = "./data/gz_2010_us_040_00_5m.json"
    )
```

# Transform Coordinate System
```{r}
us_states_aea <- spTransform(us_states,
                    CRS("+proj=laea +lat_0=45 +lon_0=-100 \
                         +x_0=0 +y_0=0 +a=6370997 +b=6370997 \
                         +units=m +no_defs"))

# Convert Row Names to Column
us_states_aea@data$id <- rownames(us_states_aea@data)
```

# Re-size and Shift States
## Alaska
```{r}
# Isolate Alaska Portion of the Shapefile
alaska <- us_states_aea[us_states_aea$STATE == "02",]

# Rotate Alaska 50 degrees
alaska <- elide(alaska, rotate=-50)

# Rescale Aaska
alaska <- elide(alaska, scale=max(apply(bbox(alaska), 1, diff)) / 2.3)

# Shift Alaska Coordiantes to Below California / Arizona
alaska <- elide(alaska, shift=c(-2.3e6, -2.475e6))

# Retrieve Projection Information
proj4string(alaska) <- proj4string(us_states_aea)
```

## Hawaii
```{r}
# Isolate Hawaii Portion of the Shapefile
hawaii <- us_states_aea[us_states_aea$STATE=="15",]

# Rotate Hawaii 35 degrees
hawaii <- elide(hawaii, rotate=-35)

# Shift Hawaii Coordiantes to Below Arizona / New Mexico / Texas
hawaii <- elide(hawaii, shift=c(5.4e6, -1.4e6))

# Retrieve Projection Information
proj4string(hawaii) <- proj4string(us_states_aea)
```

# Recombine Shifted Files
Note that Puerto Rico will be dropped in its entirety. If you have Puerto Rico data Healy recommends to shift it to between Texas and Florida
```{r}
# Drop Original Alaska, Hawaii, and Puerto Rico
us_states_aea <- us_states_aea[!us_states_aea$STATE %in% c("02", "15", "72"),]

# Bind Alaska and Hawaii
us_states_aea <- rbind(us_states_aea, alaska, hawaii)
```

# Tidy Shapefile
The united states shape-file needs to be converted into a data frame for easy use with ggplot. The State FIPS code is also created here. Once created the ID code is no longer necessary and is dropped.
```{r warning=FALSE}
us_map  <- tidy(us_states_aea, region = "GEO_ID") %>% 
  mutate(state_fips = str_sub(id,-2,-1)) %>% 
  select(-id)
```

# Export State-Level Data Frame

The final state-level data frame is saved as file *us_map.RDS*.
```{r}
saveRDS(
    object = us_map
  , file = "./us_map.RDS"
  )
```

# Inset Lines on GGPlot
I also developed lines to be places on a ggplot map to create boxes around Alaska and Hawaii. I followed the same aesthetic format as Figure 7.1 in Healy. 

## State Map Without Lines
```{r}
p <- ggplot()

p1 <- p + geom_polygon(
    data = us_map 
  , mapping = aes(x = long, y = lat, group = group, fill = state_fips)
  , color = "gray90"
  , alpha = 0.75
  , size = 0.5) + 
  coord_equal() +
  scale_fill_viridis_d() +
  guides(fill = "none") +
  theme_void()

p1
```

## State Map With Inset Lines
```{r}
# Line Properties
line_color = "black"
line_size = 0.5

p2 <- p1 + 
  geom_segment(aes(x = -2.10e6, y = -1.50e6, xend = -1.15e6, yend = -1.50e6)
               , color =  line_color, size = line_size
               ) + 
  geom_segment(aes(x = -1.15e6, y = -1.50e6, xend = -7.75e5, yend = -1.85e6)
               , color =  line_color, size = line_size
               ) +
  geom_segment(aes(x = -7.75e5, y = -1.85e6, xend = -7.75e5, yend = -2.20e6)
               , color =  line_color, size = line_size
               ) +
  geom_segment(aes(x = -7.75e5, y = -1.85e6, xend = -2.90e5, yend = -1.85e6)
               , color =  line_color, size = line_size
               ) +
  geom_segment(aes(x = -2.90e5, y = -1.85e6, xend =  1.00e5, yend = -2.20e6)
               , color =  line_color, size = line_size
               ) +
  geom_segment(aes(x =  1.00e5, y = -2.20e6, xend =  1.00e5, yend = -2.40e6)
               , color =  line_color, size = line_size
               ) +
  geom_segment(aes(x = -7.75e5, y = -2.20e6, xend = -5.60e5, yend = -2.40e6)
               , color =  line_color, size = line_size
               )

p2
```

